paper

### 2019-01-11 Report/paper.tex
push change

    \documentclass[journal,twocolumn]{IEEEtran}
\usepackage[noadjust]{cite}

@@

\begin{abstract}
    In this report we explore ways to automatically analyze ICLS and CSCL
papers to gain an understanding of the community formed by conference
participants and of the research being conducted within the community.
After describing how to extract data from the original dataset we derive
insights into the community using network analysis, natural language
processing and text mining. We discover patterns of collaboration
between participants, their migration and the global spread of the
community. We extract keywords and find clusters of referenced papers
and documents within the conferences.
\end{abstract}
    \hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

The starting point of the project are the proceedings of the
International Conference of the Learning Sciences (ICLS), respectively
the International Conference on Computer-Supported Collaborative
Learning (CSCL). The two conferences are organized by the International
Society of the Learning Sciences (ISLS) and held in biennial alternation
with each other. At the time of writing the dataset contains 4 years of
the proceedings, staring from 2015 to 2018. While ICLS covers the entire
field of the learning sciences, CSCL focuses on learning through
collaboration with the help of communication technologies \cite{ICLS}.
To understand the community formed by the participants of the
conferences as well as to understand the research being conducted within
the community, we explore methods to bring insight into the following
aspects:

\begin{itemize}
\tightlist
\item
  The institutions and countries dominant in the conferences
\item
  The differences between ICLS and CSCL with respect to citations,
  contributors and publications
\item
  First insight into changes of the years with respect to citations,
  contributors and publications
\item
  Collaboration network patterns with respect to countries, institutions
  and authors
\item
  The migration of participants across institutions
\item
  Trends of keywords over the years
\item
  Trends of citations and evolution of the number of documents from 1977
  until 2018
\item
  Clusters of the cited documents
\item
  Clusters of the papers written within the conferences
\end{itemize}

@@

    The provided raw dataset consists of papers in \emph{pdf} format, as
well as associated metadata in \emph{xml} format. To extract information
from the dataset we first convert the data to a format that is more
adapt to the task of data analysis and more compatible with a variety of
external libraries. We store all the resulting data in \emph{csv}
format, as csv is a human readable format and can easily be processed by
all the data analysis tools we employ.

@@

We extract the metadata from the xml files using the \emph{lxmltree}
python library. We associate the contents of each xml tag to a column in
csv file. As the xml files do not exactly follow the xml standard, we
handle errors by iteratively parsing the xml tree. For some files, the
section containing the keywords is malformed which can not be handled by
the parser. This is remedied by using regular expressions to match the
section contains the keywords. To each file we associate the name of the
source file and the order in which an author name occurs in the xml
file. Each author is associated one line in the XML file. From the
associated citation string we add to the first part of the citation
containing the author names, year and title. which we use to identify
references in other papers to that paper. Then we add a shortened
version of author names (i.e.~West, R.) by considering the author order.
Note that the short name extracted from the citation and the
corresponding long name (Robert West) in the metadata do at times not
match up. Thus the author order listed in the citation string does not
always correspond to the author order listed in the metadata. After
extraction, we clean the data by unifying names, as slight spelling
differences exist within the same author name (i.e.~Robert West, Robert
A. West). When unifying the names we need to make sure to not overmerge
by labeling to different authors as having the same name. First, we use
a strict condition to catch small differences in names, such as an extra
middle name. This is done by splitting up names into substrings based on
whitespace and commas. For two names in the dataset we check whether the
intersection of their sets of substrings contains at least 2 names of
length at least 3. If the condition is met we consider these two names
to be identical, and add the mapping from one name to the other to a
dictionary. Once all tuples of names (s.t. the first name appears before
the second one in order) are checked and added to the dictionary we
unify the names. Using this method we reduce the number of authors in
the dataset from 1951 down to 1879. Further, to merge names with
misspellings or people that use nicknames (Christian vs Chris) and avoid
over-merging we consider only names of authors that have similar
collaboration patterns. To do this we build the graph of co-authors and
check names of authors that are in the same neighborhood of a node on
the graph (see methodology for further details). Then we use the
\emph{difflib} library to check for similarity of two strings and
consider a name to belong to the same person if they reach the threshold
of 0.8. Again names considered to belong to the same person are added to
the mapping dictionary which is then applied to all names. This method
again reduces the number of authors in the dataset from 1879 to 1951.

@@

To extract information from the papers themselves we need to first
extract the text from the pdfs. Various parsers performing this task
exist. However, parsers such as pdfminer, PyPDF2 that claim to be able
to extract the text as well as metadata from papers have not been kept
up to date, which leads to issues with dependencies. We found that the
most reliable parser is \emph{poppler} \cite{poppler} which can be
installed and then used via the command line to extract the text
contained in the pdfs to \emph{txt} format. All methods describing how
to extract data from the papers work on the extracted \emph{txt} files.
The characters contained in the \emph{txt} files are not all encoded
uniformly, which means that characters that look the same to humans may
be encoded in different way . Hence, for many applications we need to
unify the formating. This is done using the \emph{unicodedata} module
and setting converting all characters to `Normal Form Composed' form .

A convenient property of papers is that they are structured fairly
regularly. The ISLS Author Guidelines \cite{guidelines} give us guidance
on how the papers in our dataset should be structured. We use this to
our advantage when extracting information from the paper text. While the
vast majority of papers in the dataset follow these guidelines,
exceptions do exist. We thus make sure to handle deviations from the
structure.

@@

We know that each valid paper contains a reference section at the end
listing all the references cited in the paper. While tools are being
developed to extract the reference section directly from pdfs, notably
Scholarcy \cite{scholarcy}, they did not produce satisfactory results.
That is they had issues splitting the references correctly and
categorizing parts of the reference accurately. This may occur because
they are trained to extract any reference format, and thus can not make
many assumptions on the structure. But in our dataset papers following
the guidelines should list their references in APA format. Hence using
regular expressions and background on how APA references gives us better
results than out of the box methods.

\textbf{Regular expressions in python:} Python has two modules that
implement regular expressions \emph{re} and \emph{regex}. We used regex
as --unlike the re module-- this module allows to handle strings with
non-ascii characters by having unicode expressions. As many authors have
names with non-ascii characters this is an important capability.

\textbf{To get the reference section for of a paper} we use a regular
expressions to find the reference header and determine the sections end
by checking for acknowledgement or appendix sections in the text file
previously extracted using poppler. Then, to extract individual
references we split the section into a list of substrings at each new
line character and glue substrings back together if they satisfy a set
of conditions. We found this method to be robust as it makes sure that
all properly formatted references are separated.

\textbf{To unify substrings} we exploit the fact that that APA
references all have the same initial structure: \emph{Lastname F.M.
(when published) Title which may contain colons and other non-alphabetic
chars. Where published. Who published} \cite{apaformat}. We thus check
that the beginning of each reference contains ``Lastname F.M. (when
published)''. It is relatively easy to check for this using regular
expressions as author names and year have a given structure. If a string
does not contain such a section, then it must be part of an other
reference, which is lies before in the list of reference substrings.
Note that this beginning is also what is added to the metadata as
identifying string. We found that in practical application it is
important to first merge lines if they were clearly cut off, as papers
with many authors may have the beginning on multiple lines. We can find
split up lines by checking for lines starting with non-alphabetic
characters or starting in a sentence. We merge these with the line
above. We can identify lines by their ending as being split up, such as
if they end in a comma, colon or only contain author names. We merge
these lines with the lines below them. Only then can we check whether a
line contains a valid citation start. If it does not we merge it with
the line above it. After this merging process we validate the returned
reference split by looking at the overall length and the initials, as
well as checking that a random sample returns satisfactory splitting
accuracy. We find that the above method of extracting the references
produces good results for proper APA citations and on APA citations
indeed outperforms Scolarcy when it comes to splitting up the lines.
\emph{To then get the subcomponents of a reference} is rather trivial,
as we can again use the structure of APA citations. The string before
the first parenthesis contains the authors names, inside the parentheses
we find the year or some other reference to publication time, after
which the title follows until the first dot. Then a reference to the
publishing venue follows.

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_10_0.png}\end{center}
        \caption{Different formats used to reference institution}
        \label{figA}
    \end{figure}

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_11_0.png}\end{center}
        \caption{Different formats used to reference institution}
        \label{figB}
    \end{figure}

@@

Now we consider the section before the abstract, which contains the
authors names, their email and the institution they are affiliated with.
Affiliation with institution can give us a lot of insight where people
are working and gives use a way estimate where they are from. To go from
this section to an exact affiliation with an institution a multifaceted
approach was used.

@@

The most robust way to get a persons institution affiliation is to use
their email. Emails can easily be extracted from the header section
using a regular expression and given a domain-university mapping
\cite{mapping} give us an onto mapping from domain to institution. The
additional advantage of this method is that using the additional
metadata associated to each university in the mapping, we also access
information about the country of the university.

@@

The header section itself contains information on the institution, but
as different was to list authors are accepted, notably two ways as can
be seen in Figure \ref{figA} and Figure \ref{figB}. Hence, we first
detect the structure of the header section before extracting the
association. We tag the substring present in this section using a
combination of \emph{named entity detection}, using the implementation
provided by \emph{spacy}, simple string comparison matching of the names
of authors and institutions we extracted from the metadata respectively
from Approach 1 and tagging sections containing ``emails'' using regex ,
to detect where in the header the email is. Then, taking the tagged
section, we detect which subsections correspond to which
author-affiliation listing type, and extract the author email -
institution mapping accordingly.

Approach 2 allows to match many more institutions to authors than
Approach 1 as not every institution is present in the mapping --no
companies will be in the mapping-- and many participants use gmail and
similar providers as their official correspondence email. However,
Approach 2 may contain differences in spelling in the institution name,
which need to be merged. Thus we combine both approaches, and use the
corresponding institutions in Approach 1 to unify the institutions
obtained in Approach 2. Furthermore, we later merge names by considering
the location of the institutions.

But first as with Approach 2 institution affiliation does not allow us
to automatically associate a country to the institution, we need to get
this through other means. A first step is to use the top level domains
present in the emails to associate a country of origin to the
participants and their associated institutions. Then, to plot the
community we augment the dataset with geolocation information. As a
further benefit, having this information we can then easily associate a
country to the institution for the emails that do not have a top level
domain indicating location and merge the names of institutions where
necessary based on location.

As there are a variety of way in which the name of an institution could
be written --consider EPFL, EPF Lausanne-- crawling the english, french
and german version of Wikipedia (not using the API, or the API wrapper,
as this does not give access to the desired information) was a quick and
accurate way to get the location. The advantage of sending requests over
using the API is due to the website redirecting to the relevant article
on the institution even when misspellings are present. Requests were
sent using \emph{urllib3} and \emph{beautifulsoup} was used to parse
them get the coordinates present in the article. Regretfully not all
institutions, even well known ones such as Tokyo University, have their
coordinates listed on their Wikipedia. To get the remaining locations we
use the geopy, which is a client for geocoding web services. These
results however are more error prone and will always return a result. It
is thus recommended to verify the correctness of the results. To get the
country we can then again use geopy. The resulting data containing paper
identifier, the authors email, its associated domain, order in which the
author was cited, the institution, country, Longitude and Latitude are
saved to a csv file. We use the paper identifier and author order to
associate this data with the metadata to link information about authors
with the papers they authored.

@@

\textbf{\emph{How it works}} : * It splits the text into sentences and
generates some words as candidates.\\
* Various punctuation signs will be treated as sentence boundaries. *
All words listed in the stopwords file will be treated as phrase
boundaries. This helps generate candidates that consist of one or more
non-stopwords. However ,it won't work in cases where the stopword is
part of the phrase. For example, `new' is listed in RAKE's stopword
list. This means that neither `New York' nor `New Zealand' can be ever a
keyword. * For each keyword candidate generated, the algorithm computes
the score of the pretended keyword, which is the sum of the scores for
each of its words (if it's a composed word). The words are scored
according to their frequency and the typical length of a candidate
phrase in which they appear. * The last step is ranking the scores ansd
selecting the ones with the highest score

@@

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_26_0.png}\end{center}
        \caption{Trends of keywords over the years}
        \label{fig1}
    \end{figure}

    We wanted then to know what are the trends of words for each specific
year, so we select distinct words for each year, it means that we select
for each year the words that were present for that year and never for
the others. We get the figure below. As we can see from the graph
\ref{fig2}, the most used word in the papers, more than 10\% of 2015
papers uses the word `peer', more than 6\% of the papers uses the words
`building', `embodied' in 2016. In 2017 the most frequent keyword on is
`reponsive' and in 2018 `tools' is present with more than 6\%.

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_28_0.png}\end{center}
        \caption{Distinct keywords over the years}
        \label{fig2}
    \end{figure}

@@

With the natural langage processing, we start with a naive method which
consists of clustering the documents using only TF-IDF. Each document
has a vector representation with values for each word present in the
document depending on the frequency of the word in this document
regarding its frequency in all the other documents. So if a word is a
keyword to the document and is not a common word than its value will be
more important than the other words in the vector of the document. Once
we have our vector representation, we select the max features parameter
as 2500, it means that we will only consider the top max\_features
ordered by term frequency across the corpus. On this word representation
we will run K-means.

@@

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_31_0.png}\end{center}
        \caption{Baseline model document clusters}
        \label{fig3}
    \end{figure}

@@

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_33_0.png}\end{center}
        \caption{Cluster using Average of GloVe vectors combined with K-means }
        \label{fig4}
    \end{figure}

@@

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_35_0.png}\end{center}
        \caption{Cluster using Weighted average of TF-IDF + GloVe  combined with K-means }
        \label{fig}
    \end{figure}

@@

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_41_0.png}\end{center}
        \caption{Intersection between references graph}
        \label{intersection}
    \end{figure}

@@

The second graph is more about the referenced documents and not about
the documents themselves. The aim of this graph is to see if there are
documents that are more likely to cited together. If it is the case,
then these documents are likely to be about the same subject. It is a
weighted graph constructed as follow: for each document, we create a
link between the references of the document. We create a dictionnary of
links, if the link is already in the dictionnary then we increase the
weight of the link. So for instance, if two references are present
together in the 5 distinct documents, then the weight of the edge would
be of 5. We obtain a graph of 10845 nodes and 148594 edges represented
in this figure \ref{fig6}

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_45_0.png}\end{center}
        \caption{}
        \label{}
    \end{figure}

@@

To get insight into collaborations within the community we build graphs
based on co-authorship of papers. To build the graph structure and
visualize it the python package \emph{networkx} was used. When
constructing the graph, we associate qualitative features --such as the
name and institution of the author that is represented by a node-- to
the nodes and edges of the graph.

@@

First, we built a co-authorship graph based on individual authors. Each
node represents an author, and two nodes are connected if the authors
collaborated on a paper. Other works \cite{cheong2009social},
\cite{hesford2006management} have used a directed graph that only adds a
link from the first author of the paper to all other authors. However,
as we are more interested in collaboration patterns and less in finding
the most influential first author, we build an undirected graph with
edges between all co-authors. On the downside, this yields a very dense
graph. Hence, we prune the initial graph by removing nodes that do not
satisfy additional conditions. Intuitively it is clear that a singular
collaboration on a paper does not necessarily indicate that two people
even know each other. A first condition was to only link two nodes
together if the authors corresponding to each node have appear as
co-authors in more than one conference. The graph produced by this
approach is much more sparse and contains much fewer authors.
Additionally all connected components contain fewer than 50 authors,
making it possible to consider each author individually. Such a
component is shown in figure \ref{fig_plot}.

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_51_0.png}\end{center}
        \caption{One of the connected components of the co-authorship graph}
        \label{fig_plot}
    \end{figure}

    A second alternative approach was to consider the strength of bounds. In
the graph authors are only linked if they have written at least 2 papers
together. This method yields a much larger component --containing 441
authors-- as this second condition is weaker. To get an understanding of
author to author relations in this component we need to further prune
away parts of the graph that are not insightful. When we look at
individual author within the community, we are interested well
collaborative individuals. Hence we remove nodes that have collaborated
with fewer than 3 connections to other authors. The resulting graph can
be seen in figure \ref{author_collab_large}. Details creating such a
plot can be found in the section dedicated to this below.

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_53_0.png}\end{center}
        \caption{Largest Component}
        \label{author_collab_large}
    \end{figure}

>>>>

    \documentclass[article,twocolumn]{IEEEtran}
\usepackage[noadjust]{cite}

@@

\begin{abstract}
    In this report we explore ways to automatically analyze ICLS and CSCL
papers to gain an understanding of the community formed by conference
participants and of the research being conducted within the community.
After describing how to extract data from the base dataset we derive
insights into the community using network analysis, natural language
processing and text mining. We discover patterns of collaboration
between participants, their migration and the global spread of the
community. We extract keywords and find clusters of referenced papers
and documents within the conferences.
\end{abstract}
    \hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

The starting point of the project are the proceedings of the
International Conference of the Learning Sciences (ICLS), respectively
the International Conference on Computer-Supported Collaborative
Learning (CSCL). The two conferences are organized by the International
Society of the Learning Sciences (ISLS) and held in biennial alternation
with each other. At the time of writing the dataset contains 4 years of
the proceedings, from 2015 to 2018. While ICLS covers the entire field
of the learning sciences, CSCL focuses on learning through collaboration
with the help of communication technologies \cite{ICLS}. To understand
the community formed by the participants of the conferences as well as
to understand the research being conducted within the community, we
explore methods to bring insight into the following aspects:

\begin{itemize}
\tightlist
\item
  The institutions and countries dominant in the conferences
\item
  The differences between ICLS and CSCL with respect to citations,
  research conducted and contributors
\item
  Collaboration network patterns with respect to countries, institutions
  and authors
\item
  First insight into changes over the years with respect to contributors
  and collaboration patterns
\item
  The migration of participants across institutions
\item
  Popularity of keywords over the years
\item
  Trends inregards to citated papers and the evolution of their
  prevalence from 1977 until 2018
\item
  Discovering clusters of the cited documents
\item
  Discovering clusters of the papers written within the conferences
\end{itemize}

@@

    The provided raw dataset consists of papers in \emph{pdf} format, as
well as associated metadata in \emph{xml} format. To extract information
from the dataset we first converted the data to a format that is more
adapt to the task of data analysis and more compatible with a variety of
external libraries. We stored all the resulting data in \emph{csv}
format, as csv is a human readable format and can easily be processed by
all the data analysis tools we employ.

@@

We extracted the metadata from the xml files using the \emph{lxmltree}
python library. We associated the contents of each xml tag to a column
in a csv file. As the xml files do not exactly follow the xml standard,
we handled errors by iteratively parsing the xml tree. For some files,
the section containing the keywords is malformed and can not be
extracted by the parser. This was remedied using regular expressions to
match the section containing the keywords. To each file we associated
the name of the source file and the order in which an author name occurs
in the xml file. Each author is associated to one line in the xml file.
From the associated citation string we add to the first part of the
citation containing the author names, year and title. This is later used
to identify the paper when it is referenced in an other paper. Then we
add a shortened version of author names (i.e.~West, R.) by considering
the author order. Note that it happens that the short name extracted
from the citation and the corresponding long name (Robert West) in the
metadata do not match up. Thus the author order listed in the citation
string does not always correspond to the author order listed in the
metadata. After extraction, we cleaned the data by unifying names, as
slight spelling differences exist within the same author name
(i.e.~Robert West, Robert A. West). When unifying the names we needed to
make sure not to overmerge by labeling two different authors as having
the same name. First, we used a strict condition to catch small
differences in names, such as an extra middle name. This is done by
splitting up names into substrings based on whitespaces and commas. For
two names in the dataset we checked whether the intersection of their
sets of substrings contained at least 2 names of length at least 3. If
the condition was met we considered these two names to be identical, and
added the mapping from one name to the other to a dictionary. Once all
tuples of names (s.t. the first name appears before the second one in
order) were checked and added to the dictionary we unified the names.
Using this method we reduced the number of authors in the dataset from
1951 down to 1879. Further, to merge names with misspellings or people
who use nicknames (Christian vs Chris) and to avoid overmerging we
considered only names of authors that have similar collaboration
patterns. To do this we built a graph based on co-authors relations and
checked names of authors that were in the same neighborhood of a node on
the graph (see methodology for further details). Then, we used the
\emph{difflib} library to check for similarity of two strings and
considered a name to belong to the same person if the similarity measure
reached the threshold of 0.8. Again names considered to belong to the
same person were added to the mapping dictionary which was then applied
to all names. This method again reduced the number of authors in the
dataset from 1879 to 1951.

@@

To extract information from the papers themselves we needed first to
extract the text from the pdfs. Various parsers performing this task
exist. However, parsers such as pdfminer or PyPDF2 that claim to be able
to extract the text as well as metadata from papers have not been kept
up to date, which leads to issues with dependencies. We found that the
most reliable parser is \emph{poppler} \cite{poppler} which can be
installed and then used via the command line to extract the text
contained in the pdfs to \emph{txt} format. All methods describing how
to extract data from the papers worked on the extracted \emph{txt}
files. The characters contained in the \emph{txt} files were not all
encoded uniformly, which means that characters that look the same to
humans may be encoded in different ways. Hence, for many converting all
characters to `Normal Form Composed' form.

A convenient property of papers is that they are structured fairly
regularly. The ISLS Author Guidelines \cite{guidelines} gave us guidance
on how the papers in our dataset should be structured. We used this to
our advantage when extracting information from the paper text. While the
vast majority of papers in the dataset follow these guidelines,
exceptions do exist. We thus made sure to handle deviations from the
structure.

@@

Each valid paper contains a reference section at the end listing all the
references cited in the paper. While tools are being developed to
extract the reference section directly from pdfs, notably Scholarcy
\cite{scholarcy}, they did not produce satisfactory results. That is,
they did not split the references correctly and categorized parts of the
reference inaccurately. This may occur because they are trained to
extract any reference format, and thus can not make many assumptions on
the structure. But in our dataset papers following the guidelines should
list their references in APA format. Hence using regular expressions and
background on APA referencing gave us better results than out of the box
methods.

\textbf{Regular expressions in python:} Python has two modules that
implement regular expressions, \emph{re} and \emph{regex}. We used regex
as --unlike the re module-- this module allows to handle strings with
non-ascii characters by having unicode expressions. As many authors have
names containing non-ascii characters this is an important capability.

\textbf{To get the reference section for of a paper} we used a regular
expressions to find the reference header and determine the sections end
by checking for acknowledgements or appendix sections in the text file
previously extracted using poppler. Then, to extract individual
references we split the section into a list of substrings at each new
line character and concatenated substrings back together if they
satisfied a set of conditions. We found this method to be robust as it
makes sure that all properly formatted references are separated.

\textbf{To unify substrings} we exploited the fact that APA references
all have the same initial structure: \emph{Lastname F.M. (when
published) Title which may contain colons and other non-alphabetic
chars. Where published. Who published} \cite{apaformat}. We thus checked
that the beginning of each reference contains ``Lastname F.M. (when
published)''. It is relatively easy to check for this using regular
expressions as author names and year have a given structure. If a string
does not contain such a section, then it must be part of an other
reference, which lies before in the list of reference substrings. Note
that this beginning is also what is added to the metadata as identifying
string. We found that in practical application it is important to first
merge lines if they were clearly cut off, as papers with many authors
may have the beginning on multiple lines. We could find split up lines
by checking for lines starting with non-alphabetic characters or
starting in a sentence. We merged these with the line above. We could
identify lines by their ending as being split up, such as if they end in
a comma, colon or only contain author names. We merged these lines with
the lines below them. Only then could we check whether a line contained
a valid citation start. If it did not we merged it with the line above
it. After this merging process we validated the returned reference split
by looking at the overall length and the initials, as well as by
checking that a random sample returned satisfactory splitting accuracy.
We found that the above method of extracting the references produced
good results for proper APA citations and on APA citations indeed
outperformed Scolarcy when it came to splitting up the lines.

\textbf{To then get the subcomponents of a reference} was rather
trivial, as we could again use the structure of APA citations. The
string before the first parenthesis contains the authors names, inside
the parentheses we find the year or some other reference to publication
time, after which the title follows until the first dot. Then a
reference to the publishing venue follows.

@@

Now we consider the section before the abstract, which contains the
authors names, their email addresses and the institutions they are
affiliated with. Affiliation with institution can give us a lot of
insight where people are working and gives use a way estimate where they
are from. To extract the exact affiliation of an authors with an
institution two approaches were used.

@@

The most robust way to get a persons institution affiliation is to use
their email address. Email addresses can easily be extracted from the
header section using a regular expression. Then we used a
domain-university mapping \cite{mapping} to map domains to institutions.
The additional advantage of this method is that using the additional
metadata associated to each university in the mapping, we also access
information about the country of the university.

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_11_0.png}\end{center}
        \caption{Different formats used to reference institution}
        \label{figA}
    \end{figure}

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_12_0.png}\end{center}
        \caption{Different formats used to reference institution}
        \label{figB}
    \end{figure}

@@

The header section itself contains information on the institution, but
different ways to list authors are accepted, notably two ways as can be
seen in Figure \ref{figA} and Figure \ref{figB}. Hence, we first
detected the structure of the header section before extracting the
association. We tagged the substring present in this section using a
combination of \emph{named entity detection}, using the implementation
provided by \emph{spacy}, simple string comparison matching of the names
of authors and institutions we extracted from the metadata respectively
from Approach 1 and tagging sections containing ``emails'' using regex ,
to detect where in the header the email is. Then, taking the tagged
section, we detected which subsections corresponded to which
author-affiliation listing type, and extracted the author email -
institution mapping accordingly.

Approach 2 allows to match many more institutions to authors than
Approach 1 as not every institution is present in the mapping --no
companies will be in the mapping-- and many participants use gmail and
similar providers as their official email. However, Approach 2 may
contain differences in spelling in the institution name, which needed to
be merged. Thus we combined both approaches, and used the corresponding
institutions in Approach 1 to unify the institutions obtained in
Approach 2. Furthermore, we later merged names by considering the
location of the institutions.

But first as with Approach 2 institution affiliation did not allow us to
automatically associate a country to the institution, we needed to get
this through other means. A first step was to use the top level domains
present in the emails to associate a country of origin to the
participants and their associated institutions. Then, to plot the
community we augmented the dataset with geolocation information. As a
further benefit, having this information we could then easily associate
a country to the institution for the emails that do not have a top level
domain indicating location and further merge the names of institutions
based on location.

As there are a variety of way in which the name of an institution could
be written --consider EPFL, EPF Lausanne-- crawling the english, french
and german version of Wikipedia (not using the API, or the API wrapper,
as this does not give access to the desired information) was a quick and
accurate way to get the location. The advantage of sending requests over
using the API is due to the website redirecting to the relevant article
on the institution even when misspellings are present. Requests were
sent using \emph{urllib3} and \emph{beautifulsoup} was used to parse
them get the coordinates present in the article. Regretfully not all
institutions, even well known ones such as Tokyo University, have their
coordinates listed on Wikipedia. To get the remaining locations we used
the geopy, which is a client for geocoding web services. These results
however are more error prone and will always return a result. It is thus
recommended to verify the correctness of the results. To get the country
we then again used geopy. The resulting data containing paper
identifier, the authors email, its associated domain, order in which the
author was cited, the institution, country, Longitude and Latitude were
saved to a csv file. We used the paper identifier and author order to
associate this data with the metadata to link information about authors
with the papers they authored.

@@

\textbf{\emph{How it works}} :

\begin{itemize}
\tightlist
\item
  It splits the text into sentences and generates some words as
  candidates.\\
\item
  Various punctuation signs will be treated as sentence boundaries.
\item
  All words listed in the stopwords file will be treated as phrase
  boundaries. This helps generate candidates that consist of one or more
  non-stopwords. However ,it won't work in cases where the stopword is
  part of the phrase. For example, `new' is listed in RAKE's stopword
  list. This means that neither `New York' nor `New Zealand' can be ever
  a keyword.
\item
  For each keyword candidate generated, the algorithm computes the score
  of the pretended keyword, which is the sum of the scores for each of
  its words (if it's a composed word). The words are scored according to
  their frequency and the typical length of a candidate phrase in which
  they appear.
\item
  The last step is ranking the scores ansd selecting the ones with the
  highest score
\end{itemize}

@@

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_27_0.png}\end{center}
        \caption{Trends of keywords over the years}
        \label{fig1}
    \end{figure}

    We wanted then to know what are the trends of words for each specific
year, so we select distinct words for each year. On other words it means
that we select for each year the words that were only present for this
specific year and never appeared in the others. We get the figure below.
As we can see from the graph \ref{fig2}, the most used word in the
papers, more than 10\% of 2015 papers uses the word `peer', more than
6\% of the papers uses the words `building', `embodied' in 2016. In 2017
the most frequent keyword on is `reponsive' and in 2018 `tools' is
present with more than 6\%.

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_29_0.png}\end{center}
        \caption{Distinct keywords over the years}
        \label{fig2}
    \end{figure}

@@

With the natural langage processing, we start with a naive method which
consists of clustering the documents using only TF-IDF. Each document
has a vector representation with values for each word present in the
document depending on the frequency of the word in this document
regarding its frequency in all the other documents. So if a word is a
keyword to the document and is not a common word than its value will be
more important than the other words in the vector of the document. Once
we have our vector representation, we select the max features parameter
as 2500, it means that we will only consider the top max\_features
ordered by term frequency across the corpus. On this word representation
we run K-means.

@@

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_32_0.png}\end{center}
        \caption{Baseline model document clusters}
        \label{fig3}
    \end{figure}

@@

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_34_0.png}\end{center}
        \caption{Cluster using Average of GloVe vectors combined with K-means }
        \label{fig4}
    \end{figure}

@@

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_36_0.png}\end{center}
        \caption{Cluster using Weighted average of TF-IDF + GloVe  combined with K-means }
        \label{fig}
    \end{figure}

@@

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_42_0.png}\end{center}
        \caption{Intersection between references graph}
        \label{intersection}
    \end{figure}

@@

The second graph is more about the referenced documents and not about
the documents themselves. The aim of this graph is to see if there are
documents that are more likely to cited together. If it is the case,
then these documents are likely to be about the same subject. It is a
weighted graph constructed as follow: for each document, we create a
link between the references of the document. We create a dictionnary of
links, if the link is already in the dictionnary then we increase the
weight of the link. So for instance, if two references are present
together in the 5 distinct documents, then the weight of the edge would
be of 5. We obtain a graph of 10845 nodes and 148594 edges but we
filtered the edges that have a weight higher than 4 and we get the graph
below represented in this figure \ref{fig6}

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_46_0.png}\end{center}
        \caption{Relations between references filtered}
        \label{fig6}
    \end{figure}

@@

To get insight into collaborations within the community we built graphs
based on co-authorship of papers. To build the graph structure and
visualize it the python package \emph{networkx} was used. When
constructing the graph, we associate qualitative features --such as the
name and institution of the author that is represented by a node-- to
the nodes and edges of the graph.

@@

First, we built a co-authorship graph based on individual authors. Each
node represented an author, and two nodes were connected if the authors
collaborated on a paper. Other works \cite{cheong2009social},
\cite{hesford2006management} have used a directed graph that only adds a
link from the first author of the paper to all other authors. However,
as we were more interested in collaboration patterns and less in finding
the most influential first author, we built an undirected graph with
edges between all co-authors. On the downside, it yielded a very dense
graph. Hence, we pruned the initial graph by removing nodes that did not
satisfy additional conditions. Intuitively it is clear that a singular
collaboration on a paper does not necessarily indicate that two people
even know each other. A first condition was to link two nodes together
only if the authors corresponding to each node appeared as co-authors in
more than one conference. The graph produced by this approach was much
more sparse and contained much fewer authors. Additionally all connected
components contained fewer than 50 authors, making it possible to
consider each author individually. Such a component is shown in figure
\ref{fig_plot}.

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_52_0.png}\end{center}
        \caption{One of the connected components of the co-authorship graph}
        \label{fig_plot}
    \end{figure}

    A second alternative approach was to consider the strength of bounds. In
the graph authors are only linked if they have written at least 2 papers
together. This method yielded a much larger component --containing 441
authors-- as this second condition was weaker. To get an understanding
of author to author relations in this component we needed to further
prune away parts of the graph that are not insightful. When we looked at
individual author within the community, we were interested in
collaborative individuals. Hence we removed nodes that had collaborated
with fewer than 3 connections to other authors. The resulting graph can
be seen in figure \ref{author_collab_large}. Details creating such a
plot can be found in the section dedicated to this below.

    \begin{figure}
        \begin{center}\adjustimage{max size={\linewidth}{0.4\paperheight}}{paper_files/paper_54_0.png}\end{center}
        \caption{Largest Component}
        \label{author_collab_large}
    \end{figure}

    Focusing in on the authors is not the only way to interpret the
co-authorship graph. One should also consider the overall structure of
the graph. As graph is not connected, and contains one large main
component with 1141 authors and 208 very small components with less than
25 authors in them we focused on analyzing the structure of the main
component of the graph. First we aimed to detect sub-communities within
the graph. Modularity is one measure used to find such communities. It
measures the connection strength within a community relative to the
strength of outgoing connections. Networks with high modularity have
dense connections between the nodes within modules but sparse
connections between nodes in different modules. A non-parametric
algorithm for community detection is the Louvain Method
\cite{blondel2008fast}. We used the implementation of Thomas Aynaud
\cite{louvain}. This gives a good partitioning of the graph into 26
communities of median size 42.84. This is quite a large number of
communities, so we considered an alternative way to yield bigger
partitions. An alternate way to find subcommunities is to look at
normalized cuts, which partition the nodes into groups with many
within-group connections and relatively few between-group connections.
The advantage of this method is that we can choose how many cuts to
make, and we are thus able to define the size of subcommunities. Finding
normalized cuts in a graph can be done using \emph{sklearns} spectral
clustering algorithm by passing it the adjacency matrix of the graph.
With this method we are able to neatly separate the main component graph
in three large communities. See figure \ref{normalized}
